---
title: "ECON 3201 Assignment 1"
author: "Firas Ayari"
format: typst
execute:
  echo: true
  warning: false
---

# Q1

My personal ECON 3201 Quarto website can be accessed using the following link:

**Website:** [PASTE GITHUB WEBSITE LINK HERE]

The website contains a Home page with my name, credentials, and an About Me section, as well as a Blog page for my ECON 3201 assignments.

# Q2

Let $X = \{1,2,3,\ldots,12\}$ be uniformly distributed with

$$
f_X(x)=\frac{1}{12}.
$$

## (a)

For a valid probability distribution, all probabilities must be non-negative and their sum must equal 1.

$$
\sum_{x=1}^{12} f_X(x)
=
\sum_{x=1}^{12}\frac{1}{12}
=
12\left(\frac{1}{12}\right)
=
1.
$$

Also,

$$
f_X(x)=\frac{1}{12}\geq0.
$$

Therefore, $f_X(x)$ is a valid probability distribution.

## (b)

The expected value is

$$
E(X)=\sum_{x=1}^{12}x f_X(x).
$$

Therefore,

$$
E(X)
=
\frac{1}{12}(1+2+3+\cdots+12)
=
\frac{78}{12}
=
\boxed{6.5}.
$$

## (c)

First,

$$
E(X^2)
=
\frac{1}{12}\sum_{x=1}^{12}x^2
=
\frac{650}{12}.
$$

The variance is

$$
Var(X)=E(X^2)-[E(X)]^2.
$$

Thus,

$$
Var(X)
=
\frac{650}{12}-(6.5)^2
=
\boxed{\frac{143}{12}}
\approx
\boxed{11.9167}.
$$

# Q3

Let

$$
X\sim N(4,5).
$$

Thus,

$$
\mu=4,\qquad \sigma^2=5,\qquad \sigma=\sqrt{5}.
$$

## (a)

We want

$$
P(X\leq6).
$$

Standardizing,

$$
Z=\frac{6-4}{\sqrt{5}}
=0.8944.
$$

Therefore,

$$
P(X\leq6)
=
P(Z\leq0.8944)
\approx
\boxed{0.8145}.
$$

## (b)

Since 4 is the mean of the normal distribution,

$$
P(X>4)=\boxed{0.5}.
$$

## (c)

We want

$$
P(-1\leq X<1).
$$

Standardizing both endpoints,

$$
P\left(
\frac{-1-4}{\sqrt{5}}
\leq Z <
\frac{1-4}{\sqrt{5}}
\right).
$$

Therefore,

$$
P(-2.2361\leq Z<-1.3416)
\approx
\boxed{0.0772}.
$$

# Q4

Let

$$
Z\sim N(0,1).
$$

## (a)

$$
P(Z\leq0.83)
=
\Phi(0.83)
\approx
\boxed{0.7967}.
$$

## (b)

By symmetry,

$$
P(-0.83\leq Z\leq0.83)
=
2\Phi(0.83)-1.
$$

Therefore,

$$
P(-0.83\leq Z\leq0.83)
=
2(0.7967)-1
\approx
\boxed{0.5935}.
$$

# Q5

Let $X=\{1,2,\ldots,12\}$ with $f_X(x)=1/12$.

## (a)

The probability distribution function is plotted below.

```{r}
x <- 1:12
fx <- rep(1/12, 12)

plot(
  x, fx,
  type = "h",
  lwd = 3,
  xlab = "X",
  ylab = "fX(x)",
  main = "Probability Distribution of X"
)

points(x, fx, pch = 19)
```

## (b)

The cumulative probability distribution is:

```{r}
Fx <- cumsum(fx)

plot(
  x, Fx,
  type = "s",
  xlab = "X",
  ylab = "F(X)",
  main = "Cumulative Distribution of X"
)

points(x, Fx, pch = 19)
```

## (c)

The expected value is computed in R as follows:

```{r}
EX <- sum(x * fx)
EX
```

Therefore,

$$
\boxed{E(X)=6.5}.
$$

## (d)

The variance is:

```{r}
EX2 <- sum(x^2 * fx)
variance <- EX2 - EX^2
variance
```

Therefore,

$$
\boxed{Var(X)=11.9167}.
$$

# Q6

The standard normal density is

$$
f(x)=\frac{e^{-x^2/2}}{\sqrt{2\pi}}.
$$

## (a)

The function can be defined in R as:

```{r}
f <- function(x) {
  exp(-(x^2)/2) / sqrt(2*pi)
}
```

## (b)

Integrating the function over $(-\infty,\infty)$:

```{r}
integrate(f, -Inf, Inf)$value
```

The result is

$$
\boxed{1}.
$$

Therefore, the total probability is equal to 1.

## (c)

The expected value is:

```{r}
EZ <- integrate(
  function(x) x * f(x),
  -Inf, Inf
)$value

EZ
```

Therefore,

$$
\boxed{E(Z)=0}.
$$

## (d)

The variance is:

```{r}
EZ2 <- integrate(
  function(x) x^2 * f(x),
  -Inf, Inf
)$value

VarZ <- EZ2 - EZ^2
VarZ
```

Therefore,

$$
\boxed{Var(Z)=1}.
$$

## (e)

To calculate $P(Z\leq0.83)$:

```{r}
integrate(f, -Inf, 0.83)$value
```

Therefore,

$$
\boxed{P(Z\leq0.83)\approx0.7967}.
$$

## (f)

To calculate $P(-0.83\leq Z\leq0.83)$:

```{r}
integrate(f, -0.83, 0.83)$value
```

Therefore,

$$
\boxed{P(-0.83\leq Z\leq0.83)\approx0.5935}.
$$

## (g)

Using the `pnorm()` command:

```{r}
pnorm(0.83) - pnorm(-0.83)
```

The result is approximately

$$
\boxed{0.5935}.
$$

This is the same result obtained in part (f).

# Q7

The probability density function is

$$
f(x)=\frac{1}{9}x^2,\qquad 0<x<3.
$$

## (a)

Define the function in R:

```{r}
prison_pdf <- function(x) {
  (1/9) * x^2
}
```

## (b)

Integrate over $[0,3]$:

```{r}
integrate(prison_pdf, 0, 3)$value
```

The result is

$$
\boxed{1}.
$$

Therefore, the density integrates to 1.

## (c)

The expected prison sentence is:

```{r}
expected_sentence <- integrate(
  function(x) x * prison_pdf(x),
  0, 3
)$value

expected_sentence
```

Analytically,

$$
E(X)
=
\int_0^3x\left(\frac{x^2}{9}\right)dx
=
\frac{1}{9}\int_0^3x^3dx.
$$

Thus,

$$
E(X)
=
\frac{1}{9}
\left[\frac{x^4}{4}\right]_0^3
=
\frac{81}{36}
=
\boxed{2.25}.
$$

Therefore, the expected prison sentence is **2.25 years**.

# Q8

## (a)

The standard normal distribution and Student t distributions are plotted below.

```{r}
curve(
  dnorm(x),
  from = -5,
  to = 5,
  xlim = c(-5, 5),
  ylab = "Density",
  main = "Densities of t Distribution",
  lty = 1
)

curve(
  dt(x, df = 5),
  from = -5,
  to = 5,
  col = 2,
  add = TRUE
)

curve(
  dt(x, df = 15),
  from = -5,
  to = 5,
  col = 3,
  add = TRUE
)

curve(
  dt(x, df = 25),
  from = -5,
  to = 5,
  col = 4,
  add = TRUE
)

legend(
  "topright",
  c("N(0, 1)", "M=5", "M=15", "M=25"),
  col = 1:4,
  lty = c(1, 1, 1, 1)
)
```

## (b)

As the degrees of freedom increase, the Student t distribution becomes closer to the standard normal distribution. With fewer degrees of freedom, the t distribution has heavier tails. As the degrees of freedom increase, the difference between the t distribution and the standard normal distribution becomes smaller.